2016 Presidential Election Contributions - California
=====================================================
Huey Kwik

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using
# in your analysis in this code chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk.
# This prevents the code from displaying in the knitted HTML output.
# You should set echo=FALSE for all code chunks in your file.

library(ggplot2)
library(readr)
library(dplyr)
library(Hmisc)
library(scales)
```

```{r echo=FALSE, warning=FALSE, Load_the_Data}
# Load the Data
# Warnings come from extra comma and end of each row, safe to ignore.
contributions <- read_csv("ca.csv")   
contributions$cand_nm <- factor(contributions$cand_nm)

parties <- read.csv('candidates.csv')
parties$party <- factor(parties$party, 
                        levels = c("Democratic", "Republican", "Libertarian", 
                                   "Green", "Independent"),
                        ordered = TRUE)
contributions <- contributions %>% left_join(parties, by = c("cand_nm" = "name"))
```

## Univariate Plots

### Structure of the Data

Our original dataset consists of 18 variables with over 1 million observations.
I also added party information for each candidate.

```{r echo=FALSE, Univariate_Plots}
dim(contributions)
list.tree(contributions)
```

```{r echo=FALSE, warning=FALSE, Contribution_Amounts}
# How can there be negative contributions? Clue: Redesignation to general.
ggplot(data = contributions, aes(x = contb_receipt_amt)) +
  geom_histogram(binwidth=200) +
  scale_y_log10() +
  ylab("log(count)")

summary(contributions$contb_receipt_amt)

sprintf("Number of negative contributions: %d", dim(contributions[contributions$contb_receipt_amt < 0,])[1])

#range(contributions$contb_receipt_amt, na.rm = TRUE)

# View(contributions.gt2700 <- contributions %>% 
#   filter(contb_receipt_amt > 2700))

# ggplot(data=contributions.gt2700, aes(x = contb_receipt_amt)) + 
#   geom_histogram() +
#   scale_x_continuous(breaks=seq(2000, 12000, 2000))
# 
# ggplot(data=contributions, aes(x = contb_receipt_amt)) +
#   geom_histogram() +
#   coord_cartesian(xlim = c(0,10000))

ggplot(data=contributions, aes(x = contb_receipt_amt)) + 
  geom_histogram(binwidth=200) + 
  coord_cartesian(xlim = c(0,10000)) + 
  scale_y_log10() + 
  ylab("log(count)")

contributions.gte2700 <- contributions %>%
  filter(contb_receipt_amt >= 2700) %>%
  select(cand_nm, contbr_nm, contb_receipt_amt, receipt_desc, memo_text)

ggplot(data=contributions.gte2700, aes(x = contb_receipt_amt)) + 
  geom_histogram(binwidth=200) +
  scale_x_continuous(breaks=c(2700, 5400, 8100, 10800)) +
  scale_y_log10() +
  ylab("log(count")
  

#View(contributions[is.na(contributions$contb_receipt_dt),])
#View(contributions[contributions$contb_receipt_amt < 0,])
```

Observations:

* I used log(count) so we could more easily see the data on a chart.
* Why are there negative contributions?
* Range of contributions range from -10000 to 10800
* There is one missing value for `contb_receipt_amt` in the dataset.
* Most contributions seem to be smaller
* Fewer than 1% of the contributions are negative, so it's probably okay to continue analysis and just note this.
However, still curious about why.
* Isn't there a $2700 contribution limit for individuals? We could check if contributions greater than $2700
also have negative contributions.

### Double Counting Contributions?
```{r echo=FALSE, Double_Counting_Contributions}
laguna.miguel.zip <- 92677
names <- c('HOROWITZ, DAVID', 'HOROWITZ, MICHELLE')

contributions %>% 
  filter(contbr_nm %in% names, cand_nm == "Walker, Scott") %>%
  select(contbr_nm, contb_receipt_dt, contb_receipt_amt, tran_id, election_tp)
```

I looked at some examples of contributions that were above $2700 and came across
David and Michelle Horowitz. They appear to be a couple who donated to Scott
Walker's campaign.

Summing up `contb_receipt_amt`, we get $10,800. Is this an instance of people contributing over the limit? 

From what I can tell, this instead seems to be double-counting! The FEC provides an [Individual Contributor Search](http://www.fec.gov/finance/disclosure/norindsea.shtml), which lets us look at each contributor record
in more detail. 

From there, I was able to piece this story:

* On 7/6/2015, David Horowitz donates a total of $10,800 to Scott Walker’s primary campaign.
* Of that $10,800, he reattributes $5,400 to Michelle Horowitz. 
* He redesignates $2,700 to Walker’s General Election campaign. This money gets refunded on 11/16/2015.
* The $5,400 seems appears twice, in the 3rd and 8th row of the table.
* Assuming that those records are duplicates, then Michelle Horowitz reattributes $2,700 to Walker's General Election campaign.

If this story is true, then these donations are within the campaign contribution limits for primary and
general elections. From an election integrity standpoint, this is good. 

However, when doing analysis of this data, we should be aware of this discrepancy in our analysis.
A contribution like Michelle Horowitz's reattributed $5,400 may be double-counted in our analysis. Also,
a large contribution of $10,800 by David Horowitz will count towards calculating the mean, even though
it gets reattributed into smaller contributions later.

### Party
```{r echo=FALSE, Party}
ggplot(data = contributions, aes(x = party)) + geom_bar(stat = 'count')
```

Democrats had the most contributions by far.

### Misc. Stats
```{r echo=FALSE, Misc_Stats}
range(contributions$contb_receipt_dt, na.rm = TRUE)
length(unique(contributions$contbr_employer))
length(unique(contributions$contbr_zip))
length(unique(contributions$contbr_city))
length(unique(contributions$contbr_nm))
```

* Date range: April 1, 2015 to October 31, 2016
* 60587 unique employers
* 136234 unique zipcodes
* 2418 unique cities
* 209688 unique contributors 


# Univariate Analysis

### What is the structure of your dataset?

### What is/are the main feature(s) of interest in your dataset?

### What other features in the dataset do you think will help support your investigation into your feature(s) of interest?

### Did you create any new variables from existing variables in the dataset?
Party.

### Of the features you investigated, were there any unusual distributions? Did you perform any operations on the data to tidy, adjust, or change the form of the data? If so, why did you do this?


TODOs:

* Visualize total contributions for candidate, mean, median
* See contributions over time per candidate
* Do contribution sizes differ for differing candidates? Like is Bernie Sanders the candidate of the 99%?
* Look at distribution of contributions for candidates... box plot


# Bivariate Plots Section

## Contributions per Candidate
```{r echo=FALSE, Aggregate_Contributions}
contributions.candidates <- contributions %>% 
  filter(contb_receipt_amt > 0) %>%
  group_by(cand_nm) %>%
  summarise(total = sum(contb_receipt_amt, na.rm = TRUE),
            mean = mean(contb_receipt_amt, na.rm = TRUE),
            median = median(as.numeric(contb_receipt_amt), na.rm = TRUE),
            n = n(),
            party = first(party)) %>%
  arrange(-n) %>%
  mutate(cand_nm = factor(cand_nm, cand_nm)) 

cand.order = levels(contributions.candidates$cand_nm)

CreatePlot <- function(df, x, y, ylabel) {
  df <- df %>% arrange_(y) 
    
  df[[x]] = factor(df[[x]], df[[x]], ordered = TRUE)

  p <- ggplot(data = df, aes_string(x = x, y = y)) + 
    geom_bar(stat = "identity") +
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) + 
    ylab(label = ylabel) +
    scale_y_continuous(labels = comma) + 
    coord_flip() 
  p
}

CreatePlot(contributions.candidates, "cand_nm", "n", "Number of Contributions")
CreatePlot(contributions.candidates, "cand_nm", "total", "Total Contributions, Dollars")
CreatePlot(contributions.candidates, "cand_nm", "mean", "Average Contribution, Dollars")
CreatePlot(contributions.candidates, "cand_nm", "median", "Median Contribution, Dollars")

# Bit surprised that HRC's median contribution is about the same as Bernie's. 
# Look at raw data:

summary(contributions %>% 
  filter(cand_nm == "Sanders, Bernard") %>%
  select(contb_receipt_amt))

summary(contributions %>% 
  filter(cand_nm == "Clinton, Hillary Rodham") %>%
  select(contb_receipt_amt))

levels(contributions$cand_nm) = rev(cand.order)

ggplot(data = contributions, aes(x = cand_nm, y = contb_receipt_amt)) + 
  geom_boxplot() + coord_flip()

ggplot(data = subset(contributions, contributions$contb_receipt_amt > 0), 
       aes(x = cand_nm, y = contb_receipt_amt)) + 
  geom_boxplot() + coord_flip()
```

I was a bit surprised that Clinton's median donation was actually lower than Sanders's.

For the box plots, I sorted the candidates from higher number of donations to lowest.
For Republicans and Democrats, that roughly seems to correspond with how long they stayed in the race, 
but I'll visualize that later.

In the first box plot, we can see that some candidates actually have many donations
above the individual limit of $2700. Many also have negative donations, which could
either be refunds or reattributions. 

Some candidates did better at staying within contribution limits than others. For 
instance, Clinton did not have a single negative contribution or a contribution that
exceeded the individual limit. The same could be said for Ben Carson and Jeb Bush.


# Bivariate Analysis

### Talk about some of the relationships you observed in this part of the investigation. How did the feature(s) of interest vary with other features in the dataset?

### Did you observe any interesting relationships between the other features (not the main feature(s) of interest)?

### What was the strongest relationship you found?




# Multivariate Plots Section

```{r echo=FALSE, Multivariate_Plots}

```

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the investigation. Were there features that strengthened each other in terms of looking at your feature(s) of interest?

### Were there any interesting or surprising interactions between features?

### OPTIONAL: Did you create any models with your dataset? Discuss the strengths and limitations of your model.

------

# Final Plots and Summary

### Plot One
```{r echo=FALSE, Plot_One}

```

### Description One


### Plot Two
```{r echo=FALSE, Plot_Two}

```

### Description Two


### Plot Three
```{r echo=FALSE, Plot_Three}

```

### Description Three

------

# Reflection
