2016 Presidential Election Contributions for California  by Huey Kwik
========================================================

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using
# in your analysis in this code chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk.
# This prevents the code from displaying in the knitted HTML output.
# You should set echo=FALSE for all code chunks in your file.

library(ggplot2)
library(readr)
library(dplyr)
```

```{r echo=FALSE, warning=FALSE, Load_the_Data}
# Load the Data
# Warnings come from extra comma and end of each row, safe to ignore.
contributions <- read_csv("ca.csv")   
contributions$cand_nm <- factor(contributions$cand_nm)

parties <- read.csv('candidates.csv')
parties$party <- factor(parties$party, 
                        levels = c("Democratic", "Republican", "Libertarian", 
                                   "Green", "Independent"),
                        ordered = TRUE)
contributions <- contributions %>% left_join(parties, by = c("cand_nm" = "name"))
```

# Univariate Plots Section

Our dataset consists of 18 variables with over 1 million observations.

```{r echo=FALSE, Univariate_Plots}
dim(contributions)
str(contributions)
```

```{r echo=FALSE, warning=FALSE, Contribution_Amounts}
# How can there be negative contributions? Clue: Redesignation to general.
ggplot(data = contributions, aes(x = contb_receipt_amt)) +
  geom_histogram()

summary(contributions$contb_receipt_amt)

contributions.gt2700 <- contributions %>% 
  filter(contb_receipt_amt >= 2700) %>%
  select(cand_nm, contbr_nm, contb_receipt_amt, receipt_desc, memo_text)
#range(contributions$contb_receipt_amt, na.rm = TRUE)

# View(contributions.gt2700 <- contributions %>% 
#   filter(contb_receipt_amt > 2700))

ggplot(data=contributions.gt2700, aes(x = contb_receipt_amt)) + 
  geom_histogram() +
  scale_x_continuous(breaks=seq(2000, 12000, 2000))

ggplot(data=contributions, aes(x = contb_receipt_amt)) + 
  geom_histogram() + 
  coord_cartesian(xlim = c(0,10000))

ggplot(data=contributions, aes(x = contb_receipt_amt)) + 
  geom_histogram() + 
  coord_cartesian(xlim = c(0,10000)) + 
  scale_y_log10()

#View(contributions[is.na(contributions$contb_receipt_dt),])

dim(contributions[contributions$contb_receipt_amt < 0,])

#View(contributions[contributions$contb_receipt_amt < 0,])
```

Misc observations:

* Why are there negative contributions?
* Range of contributions range from -10000 to 10800
* There are actually NAs for `contb_receipt_amt`
* Most contributions seem to be smaller
* Fewer than 1% of the contributions are negative, so it's probably okay to ignore for analysis.
However, still curious about why.
* Isn't there a $2700 limit per candidate? 
  * Try joining contributors who are over the limit with the rest of dataset, see if they get refunds

```{r echo=FALSE, Double_Counting_Contributions}
laguna.miguel.zip <- 92677
names <- c('HOROWITZ, DAVID', 'HOROWITZ, MICHELLE')

contributions %>% 
  filter(contbr_nm %in% names, cand_nm == "Walker, Scott") %>%
  select(contbr_nm, contb_receipt_dt, contb_receipt_amt, tran_id, election_tp)
```

Summing up `contb_receipt_amt` seems to be double-counting! I asked the FEC about
this, and they pointed me to the [Individual Contributor Search](http://www.fec.gov/finance/disclosure/norindsea.shtml).
This interface allows users to look at each contributor record in more detail.

From there, I was able to piece this story:

* On 7/6/2015, David Horowitz donates a total of $10,800 to Scott Walker’s primary campaign.
* Of that $10,800, he reattributes $5,400 to Michelle Horowitz. 
* He redesignates $2,700 to Walker’s General Election campaign. This money gets refunded on 11/16/2015.
* The $5,400 seems appears twice, in the 3rd and 8th row of the table.
* Assuming that those records are duplicates, then Michelle Horowitz reattributes $2,700 to Walker's General Election campaign.

If this story is true, then these donations are within the campaign contribution limits for primary and
general elections. From an election integrity standpoint, this is good. However, when doing aggregate
analysis of this data, we should be aware that some of these contributions may be double counted.

```{r echo=FALSE, Party}
ggplot(data = contributions, aes(x = party)) + geom_bar(stat = 'count')
```

Democrats had the most contributions by far.

```{r echo=FALSE, Misc_Stats}
range(contributions$contb_receipt_dt, na.rm = TRUE)
length(unique(contributions$contbr_employer))
length(unique(contributions$contbr_zip))
length(unique(contributions$contbr_city))
length(unique(contributions$contbr_nm))
```

* Date range: April 1, 2015 to October 31, 2016
* 60587 unique employers
* 136234 unique zipcodes
* 2418 unique cities
* 209688 unique contributors 

TODOs:

(/) Add Party as a column, since I'll want to look at that
* See contributions over time per candidate
* Do contribution sizes differ for differing candidates? Like is Bernie Sanders the candidate of the 99%?
(/) What's the range of dates?

# Univariate Analysis

### What is the structure of your dataset?

### What is/are the main feature(s) of interest in your dataset?

### What other features in the dataset do you think will help support your investigation into your feature(s) of interest?

### Did you create any new variables from existing variables in the dataset?
Party.

### Of the features you investigated, were there any unusual distributions? Did you perform any operations on the data to tidy, adjust, or change the form of the data? If so, why did you do this?



# Bivariate Plots Section
```{r echo=FALSE, Bivariate_Plots}

```

# Bivariate Analysis

### Talk about some of the relationships you observed in this part of the investigation. How did the feature(s) of interest vary with other features in the dataset?

### Did you observe any interesting relationships between the other features (not the main feature(s) of interest)?

### What was the strongest relationship you found?




# Multivariate Plots Section

```{r echo=FALSE, Multivariate_Plots}

```

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the investigation. Were there features that strengthened each other in terms of looking at your feature(s) of interest?

### Were there any interesting or surprising interactions between features?

### OPTIONAL: Did you create any models with your dataset? Discuss the strengths and limitations of your model.

------

# Final Plots and Summary

### Plot One
```{r echo=FALSE, Plot_One}

```

### Description One


### Plot Two
```{r echo=FALSE, Plot_Two}

```

### Description Two


### Plot Three
```{r echo=FALSE, Plot_Three}

```

### Description Three

------

# Reflection
